---
title: Expressing size in bananas a dive into {vctrs}
author: Roel M. Hogervorst
date: '2020-06-08'
categories:
  - blog
  - R
tags:
  - banana
  - package
  - intermediate
  - units
  - vctrs
  - S3
slug: expressing-size-in-bananas-a-dive-into-vctrs
subtitle: 'Yes I made a stupid package to express lengths in bananas'
output:
  html_document:
    keep_md: yes
---

<!-- useful settings for rmarkdown-->

```{r setup, include=FALSE}
# Options to have images saved in the post folder
# And to disable symbols before output
knitr::opts_chunk$set(fig.path = "", comment = "")

# knitr hook to make images output use Hugo options
knitr::knit_hooks$set(
  plot = function(x, options) {
    hugoopts <- options$hugoopts
    paste0(
      "{{<figure src=",
      '"', x, '" ',
      if (!is.null(hugoopts)) {
        glue::glue_collapse(
          glue::glue('{names(hugoopts)}="{hugoopts}"'),
          sep = " "
        )
      },
      ">}}\n"
    )
  }
)

# knitr hook to use Hugo highlighting options
knitr::knit_hooks$set(
  source = function(x, options) {
  hlopts <- options$hlopts
    paste0(
      "```r ",
      if (!is.null(hlopts)) {
      paste0("{",
        glue::glue_collapse(
          glue::glue('{names(hlopts)}={hlopts}'),
          sep = ","
        ), "}"
        )
      },
      "\n", glue::glue_collapse(x, sep = "\n"), "\n```\n"
    )
  }
)
```

<!-- content --> 
Recently I've become interested in relative sizes of things. Maybe I'm paying more attention
to my surroundings since I'm locked at home for so long. Maybe my inner child is finally
breaking free. Whatever the reason, I channeled all of that into two packages: 
- [everydaysizes](https://github.com/RMHogervorst/everydaysizes) A rather unfinished collection of dimensions of everyday objects.
- [banana](https://github.com/RMHogervorst/banana) A package that displays dimensions as ... bananas.

I've collected a bunch of sizes and turned them into 'units'. The units package in R can automatically convert units to other units. It's amazing, it can do mph to m/s and many other things. But in this
case I'm using length units: inches, centimeter and meters. 


```{r}
library(tidyverse)
library(units)
# Banana for banana vectors, and everydaysizes for the collection of metrics.
library(banana)  # https://github.com/RMHogervorst/banana
library(everydaysizes) # https://github.com/RMHogervorst/everydaysizes
```

I'm selecting two interesting sizes from the everydaysizes package: The Eiffel Tower (Tour de Eiffel in local language) and a tree I visited in New Zealand: 'the Father of the forest' (or E Matua Ngahere in the local language). 


```{r}
how_big <- 
  everydaysizes::things %>% 
  slice(1,7)
how_big %>% select(what, height)
```

So now we know how high the Eiffel Tower in Paris and the 'Father of the Forest' tree are in meters.

Since this is size is represented in 'units' it is rather trivial to turn them into imperial metrics too:

```{r}
how_big %>% 
  mutate(feet = set_units(height, "ft")) %>% 
  select(what, height, feet)
```
However, this doesn't really mean anything to me. How about in bananas?

```{r}
how_big %>% 
  mutate(feet = set_units(height, "ft")) %>% 
  mutate(bananas = as_banana(height)) %>% 
  select(what, height, feet, bananas)
```
Yes I made a stupid package to express things in bananas. Not one size though, it tries to match
it to small or large bananas and it rounds it up to nearest banana. 

What about the height and length of an Alpaca?

```{r}
Alpaca <- 
  everyday_items %>% 
  slice(6) %>% 
  select(what, height, width)
Alpaca
```

```{r}
Alpaca_banana <- 
  Alpaca %>% 
  mutate(
    banana_h = as_banana(height),
    banana_w = as_banana(width)
    )
Alpaca_banana
```

That is right. An alpaca is approximately 4 large bananas high up to the shoulder and 7 small bananas long. 


### A small dive into banana properties
I'm using *(probably the word here is 'abusing')* the vctrs package to create a new class of s3-vector. In the package I tell work out how transformations from numeric, and units and integers go to bananas and back. I made a mistake somewhere because I can multiply a number with a banana, but not a banana with a numeric. Anyways the {vctrs} package is super fun and a joy to work with! What I did was, I define a different print method for the banana vector. And so inside data.frames and the `str()` function, it uses this print method.

```{r}
bananavec <- Alpaca_banana$banana_w[[1]]
class(bananavec)
str(bananavec)
```

But since bananas is a relative metric and always in integers some behavior is a bit ... 'strange':

```{r}
bananavec2 <- c(1, 1.1, 1.3, 1.5) * bananavec
bananavec2
```

```{r}
data.frame(bananavec2, increase = c(1, 1.1, 1.3, 1.5))
```

So we start with 7 small bananas, and an increase of 10% leads to 6 large bananas!

This is because bananavectors are numeric underneath:

```{r}
as.numeric(bananavec2)
```

Or actually in centimeters:

```{r}
banana::to_units(bananavec2)
```

That is all for today. I might spend some time on the inner workings of banana later. 



### State of the machine
<details>
<summary> At the moment of creation (when I knitted this document ) this was the state of my machine: **click here to expand** </summary>

```{r}
sessioninfo::session_info()
```

</details>


