---
title: New Package, Pinboardr
author: Roel M. Hogervorst
date: '2020-05-11'
slug: new-package-pinboardr
categories:
  - R
tags:
  - package
  - bookmarks
  - beginner
  - pinboardr
subtitle: ''
share_img: 'header_image.png'
output:
  html_document:
    keep_md: yes
---

<!-- useful settings for rmarkdown-->

```{r setup, include=FALSE}
# Options to have images saved in the post folder
# And to disable symbols before output
knitr::opts_chunk$set(fig.path = "", comment = "")

# knitr hook to make images output use Hugo options
knitr::knit_hooks$set(
  plot = function(x, options) {
    hugoopts <- options$hugoopts
    paste0(
      "{{<figure src=",
      '"', x, '" ',
      if (!is.null(hugoopts)) {
        glue::glue_collapse(
          glue::glue('{names(hugoopts)}="{hugoopts}"'),
          sep = " "
        )
      },
      ">}}\n"
    )
  }
)

# knitr hook to use Hugo highlighting options
knitr::knit_hooks$set(
  source = function(x, options) {
  hlopts <- options$hlopts
    paste0(
      "```r ",
      if (!is.null(hlopts)) {
      paste0("{",
        glue::glue_collapse(
          glue::glue('{names(hlopts)}={hlopts}'),
          sep = ","
        ), "}"
        )
      },
      "\n", glue::glue_collapse(x, sep = "\n"), "\n```\n"
    )
  }
)
```

<!-- content  -->

I've created a new package to interact with [pinboard](https://pinboard.in) *not to be confused with pinterest*. I noticed there wasn't a package yet and the API is fairly clear. 
So come and check it out {pinboardr} at <https://github.com/RMHogervorst/pinboardr>

I did see a new package to interact with pocket: [pocketapi](https://github.com/CorrelAid/pocketapi).
Since pocket is also a kind of bookmark manager I thought there was a need for these kinds of
packages. I will leave this package on github for a while, to figure out if I need to
make changes and in a month or so I will push it to CRAN. Everyone who uses pinboard and R is
invited to 'kick the tyres' as they say. Let me know if the docs are clear, if it is usable
and useful for you.

### What is pinboard? 
Pinboard is a bookmark website. In stead of bookmarking on your own computer in the browser you
can use pinboard. The service is very minimalistic and privacy focused. I use it all the time
to put articles on my reading list, to archive tweets and to collect material for my secondary blog
and for this blog.

## Basic functionality
I added a more extensive readme, and all the functions are documented too. But in this
short blogpost I wanted to show some examples.

With {pinboardr} you can see what you and other people used as tags for a page:

```{r}
library(pinboardr)
caption <- "package {pinboardr}"
pb_posts_suggest(url="https://blog.rmhogervorst.nl/blog/2020/04/08/scraping-gdpr-fines/")
pb_posts_suggest("http://neverssl.com")
```


Get recent bookmarks with a specific type

```{r,message=FALSE}
library(tidyverse)
```

```{r}
recent <- pb_posts_recent(tags = "r",count=4)
names(recent)
recent %>% select(title, tags)
```

Find out when you bookmarked things. 
(I think many bookmarks have the same date because I imported and exported from one tool to another)

```{r}
dates <- pb_posts_dates()
dates %>% 
  mutate(dates= as.POSIXct(date)) %>% 
  ggplot(aes(dates, count)) +
  geom_col()+
  labs(title="bookmarks per date", caption=caption)
```

Can you guess when I was on the other side of the world with very spotty internet?


```{r}
dates %>% 
  mutate(dates= as.POSIXct(date)) %>% 
  filter(dates > "2019-08-01") %>% 
  ggplot(aes(dates, count)) +
  geom_col()+
  labs(title="bookmarks per date", subtitle = "zoomed in", caption=caption)
```

So, see if you like it and check it out at [github](https://github.com/RMHogervorst/pinboardr)

### Reproducibility
<details>
<summary> At the moment of creation (when I knitted this document ) this was the state of my machine: **click here to expand** </summary>

```{r}
sessioninfo::session_info()
```

</details>


